Optimal (Multiway) Spatial Joins

### 2024-06-03 msj.tex
yufei

\documentclass[sigconf]{acmart}
% \documentclass{...}
%\input{./def/yf-formatting}
\input{./def/yf-def}

@@

\def\vgap{\vspace{0mm}}
\def\extraspacing{\vspace{1.5mm} \noindent}
\def\figcapup{\vspace{-2mm}}
\def\figcapdown{\vspace{-3mm}}

@@

    \vspace{2mm}
    \figcapup 
    \caption{Result comparison on $k$-SJ problem for a constant $k$}
    \label{tab:results-com}
    \figcapdown \vspace{-5mm}
\end{table}

@@

When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised; the space consumption of our algorithm is $O(n + \out)$. Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary. Interestingly, the answer is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory. Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O((n + \out) \cdot \polylog n)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}. This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result. We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.

\begin{figure*}
    \begin{tabular}{ccccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=27mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-d} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
        (a) Problems $\mathscr{A}$ &
        \hspace{3mm}
        (b) Problem $\mathscr{B}$ &
        \hspace{3mm}
        (c) Problem $\mathscr{C}$ &
        \hspace{3mm}
        (d) Problem $\mathscr{D}$ &
        \hspace{3mm}
        (e) Problem $\mathscr{E}$
    \end{tabular}

\section{Preliminaries in Geometry} \label{sec:bricks}

@@

@@

\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently. They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported. Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time. We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$. Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.

\begin{figure*}
    \begin{tabular}{cccc}
        \hspace{-5mm}
        \includegraphics[height=45mm]{./artwork/alg-ex1-a} &
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex1-b} & 
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex1-c} &
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex1-d} \\ 
        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
        \hspace{-4mm} (b) $R_1'$ and $R_2'$ &
        \hspace{-4mm} (c) $R_1'$, $R_2'$, $H$, and $V$ &
        \hspace{-4mm} (d) $R_1^*$, $R_2^*$ ($= \emptyset$), and $H$
    \end{tabular}
    \figcapup 
    \caption{Finding H-V $k$-SJ result tuples of type 1 ($k = 4$)} 
    \label{fig:hv:type1:ex}
    \figcapdown 
\end{figure*}

@@

%\yf{example}

@@

\begin{figure*}
    \begin{tabular}{cccc}
        \hspace{-5mm}
        \includegraphics[height=45mm]{./artwork/alg-ex2-a} &
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex2-b} &
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex2-c} &
        \hspace{-4mm}
        \includegraphics[height=45mm]{./artwork/alg-ex2-d} \\
        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
        \hspace{-4mm} (b) $R_1$, $H'$, and $V$ &
        \hspace{-4mm} (c) $R_2$, $H'$, and $V$ &
        \hspace{-4mm} (d) $R_2$ and $H^*$
    \end{tabular}
    \figcapup
    \caption{Finding H-V $k$-SJ result tuples of type 2 ($k = 4$)}
    \label{fig:hv:type2:ex}
    \figcapdown
\end{figure*}

>>>>

%\documentclass[sigconf]{acmart}
\documentclass[acmsmall,review,anonymous]{acmart}
% \documentclass{...}
%\input{./def/yf-formatting}
\input{./def/yf-def}

@@

\def\vgap{\vspace{1mm}}
\def\extraspacing{\vspace{3mm} \noindent}
\def\figcapup{\vspace{-2mm}}
\def\figcapdown{\vspace{-3mm}}

@@

    \vspace{4mm}
    \figcapup 
    \caption{Result comparison on $k$-SJ problem for a constant $k$}
    \label{tab:results-com}
    \figcapdown \vspace{-5mm}
\end{table}

@@

When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised; the space consumption of our algorithm is $O(n + \out)$. Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary. Interestingly, the answer is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory. Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O((n + \out) \cdot \polylog n)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}. This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result. We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.
\section{Preliminaries in Geometry} \label{sec:bricks}

@@

\begin{figure*}
    \begin{tabular}{cc}
        \includegraphics[height=30mm]{./artwork/prob-a} &
        \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
        (a) Problems $\mathscr{A}$ &
        \hspace{3mm}
        (b) Problem $\mathscr{B}$ \\[3mm]
    \end{tabular}
    %\vspace{3mm}
    \begin{tabular}{ccc}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \includegraphics[height=30mm]{./artwork/prob-d} &
        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
        (c) Problem $\mathscr{C}$ &
        (d) Problem $\mathscr{D}$ &
        (e) Problem $\mathscr{E}$
    \end{tabular}

@@

\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently. They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported. Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time. Another structure with the same guarantees can be found in \cite{oa19}. We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$. Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.

@@

\begin{figure*}
    \begin{tabular}{cc}
        %\hspace{-5mm}
        \includegraphics[height=48mm]{./artwork/alg-ex1-a} &
        %\hspace{-4mm}
        \includegraphics[height=48mm]{./artwork/alg-ex1-b} \\
        %\hspace{-5mm}
        (a) $R_1$, $R_2$, $H$, and $V$ &
        %\hspace{-4mm}
        (b) $R_1'$ and $R_2'$ \\[2mm]
        %\hspace{-4mm}
        \includegraphics[height=48mm]{./artwork/alg-ex1-c} &
        %\hspace{-4mm}
        \includegraphics[height=48mm]{./artwork/alg-ex1-d} \\
        (c) $R_1'$, $R_2'$, $H$, and $V$ &
        (d) $R_1^*$, $R_2^*$ ($= \emptyset$), and $H$
    \end{tabular}
    \figcapup
    \caption{Finding H-V $k$-SJ result tuples of type 1 ($k = 4$)}
    \label{fig:hv:type1:ex}
    \figcapdown
\end{figure*}

@@

\begin{figure*}
    \begin{tabular}{cc}
        %\hspace{-5mm}
        \includegraphics[height=48mm]{./artwork/alg-ex2-a} &
        %\hspace{-4mm}
        \includegraphics[height=48mm]{./artwork/alg-ex2-b} \\
        %\hspace{-5mm}
        (a) $R_1$, $R_2$, $H$, and $V$ &
        %\hspace{-4mm}
        (b) $R_1$, $H'$, and $V$ \\ [2mm]
        \includegraphics[height=48mm]{./artwork/alg-ex2-c} &
        \includegraphics[height=48mm]{./artwork/alg-ex2-d} \\
        %\hspace{-4mm}
        (c) $R_2$, $H'$, and $V$ &
        %\hspace{-4mm}
        (d) $R_2$ and $H^*$
    \end{tabular}
    \figcapup
    \caption{Finding H-V $k$-SJ result tuples of type 2 ($k = 4$)}
    \label{fig:hv:type2:ex}
    \figcapdown
\end{figure*}
